Playing against Nature: causal discovery for decision making under uncertainty

### 2018-05-17 Segundo_Semestre/CausalML_Workshop/Paper_long_version.tex
cambios redaccion en intro y related work

\begin{abstract}
We consider decision problems under uncertainty $(A,E,C,u)$ where elements in $C$ and elements in $E$ are related through a causal mechanism which is unknown to the decision maker. 
We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge in each round.
%Interaction between the decision maker and his environment is modelled as a repeated game where a player called Nature selects his actions from the causal model.
%The decision maker has beliefs over the causal model which she updates in each round. 
\end{abstract}

\section{Introduction}
A fundamental part of intelligent reasoning  is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}). In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}). On the other hand, if enough information is not available, the decision maker could try to \textit{learn} from the environment by interacting with it.

Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely reactive and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process (\cite{garnelo2016towards}).

Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} unknown to the agent.

@@

While the standard setting in RL is to model the interaction agent-environment as an agent moving from one \textit{state} to another and observing a reward, we propose to model it as a \textit{game} between the decision maker and a player called Nature which will select his actions from the causal model.

\section{Related Work}
One stage decision problems where the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds. In a classic bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.

@@

By considering a causal model which is partially known and intervening variables from the unknown part and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.

The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices but the problem of \textit{acquiring} this causal knowledge is left unatacked. Here we propose to acquire, and use, causal information while interacting with the environment. 
\section{Problem setup}
By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).

@@

In fact, noting that the variable to be intervened by the decision maker doesn't has parents, one can explicitly calculate the probability of the target variable given a value for the treatment.

Proceeding in this way, the probability of the patient living if the assigned treatment is surgery is only $0.26$.

On the other hand, if the treatment assigned is the pill, then the living probability is $0.4$
\subsection{Only the structure is known}
If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) of the model which she must learn by repetitions.

>>>>

\begin{abstract}
We consider decision problems under uncertainty $(A,E,C,u)$ where elements in $C$ and elements in $E$ are related through a causal mechanism which is unknown to the decision maker. 
We study how a decision maker can learn about this causal mechanism through sequential decision making as well as using current causal knowledge inside each round in order to make better choices had she not considered causal knowledge.
\end{abstract}

\section{Introduction}
A fundamental part of intelligent reasoning is being able to make decisions under uncertain conditions (\cite{lake2017building}, \cite{danks2014unifying}, \cite{pearlwhy}). In some cases, a decision maker who faces an uncertain environment has enough information to make choices by maximizing expected utility, which is the formal criteria for making decisions if rational preferences are assumed (\cite{bernardo2000bayesian}, \cite{gilboa2009decision}). On the other hand, if enough information is not available, the decision maker could attempt to \textit{learn} from the environment by interacting with it.

Learning by interaction has been extensively studied by computer scientists using the Reinforcement Learning (RL) setting \cite{sutton1998reinforcement}, but the most common used techniques  in this field are purely associative and do not consider any high-level structure of the environment beyond what is expressable in a Markov Decision Process \cite{garnelo2016towards}.

Since human beings are known to learn causal graphical models in sequential decision making processes (\cite{sloman2006causal}, \cite{nichols2007decision}, \cite{meder2010observing}, \cite{hagmayer2013repeated}, \cite{danks2014unifying}), although this learning is not perfect (\cite{rottman2014reasoning}), we propose that an autonomous agent can learn and use causal information while interacting with an uncertain environment which is governed by a fixed \textit{causal mechanism} which is unknown to the agent.

@@

While the standard setting in RL is to model the interaction agent-environment as an agent moving from one \textit{state} to another inside a model of the environemtn and observing a reward as this transitions occur, we propose to model it as a \textit{game} between the decision maker and a player called Nature which will select his actions from the causal model.

\section{Related Work}
Decision problems where the actions available to a decision maker are interventions over a known causal model are analyzed by \cite{lattimoreNIPS2016} as a \textit{bandit problem} where the optimal action must be learned over $T$ rounds of action-observation in which only one action can be chosen. In a classic bandit problem an agent chooses an \textit{arm} from a slot machine, observes a reward and then moves on to the next machine which is of the same kind and whose initial settings are independent of the previous machine and action \cite{sutton1998reinforcement}.

@@

By considering a causal model which is partially known and intervening variables from the unknown part of the model and by avoiding sampling arms that are considered sub-optimal \cite{sen2017identifying} extend the work of \cite{lattimoreNIPS2016}.

The aforementioned papers assume the causal model is known to the decision makers so their work focuses on \textit{using} causal information to make good choices but the problem of \textit{acquiring} this causal knowledge is left unatacked.

In this work we propose to acquire, by repeated interaction, causal information about the environment as well as using the current causal knowledge inside each round to make better decisions. By modelling the environment as a player in a game we allow it to have objectives to pursue which will allow to model a rich family of situations where several agents are competing against each other and a causal entity controls the outcomes. 
\section{Problem setup}
By \textit{causality} we mean a stochastic binary relation between events of a probability space $(\Omega, \mathcal{F}, \mathbb{P})$ denoted by $“\to”$ that is transitive, irreflexive and antisymmetric (\cite{spirtes2000causation}).

@@

In fact, noting that the variable to be intervened by the decision maker doesn't has parents in the graph one can explicitly calculate the probability of the target variable given a value for the treatment using variable elimination (\cite{koller2009probabilistic}) given the treatment as evidence.

Proceeding in this way, the probability of the patient living if the assigned treatment is surgery is only $0.28$.

On the other hand, if the treatment assigned is the pill, then the living probability is $0.37$
\subsection{Only the structure is known}
If only the structure of the graph is known, the decision maker will have \textit{beliefs} over the parameters (conditional probabilities) of the model which she must learn by repetitions.
